Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-07 foreground_delay_spectrum_modeling.tex
updated data processing, clean and leakage sections

\subsubsection{Initial Data processing}\label{sec:data-analysis}
The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference, removing 3\% of the data and averaged in time  and frequency from the raw 1/2s,40kHz to 2s, 80khz. These data are then calibrated to a simulation of the sky containing 2420 point sources selected from the MWA Commissioning Survey (N.H Walker, in prep).  MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.        The calibration sources are selected to fall within the 10\% primary lobe of the tile beam. The calibration algorithm computes a per channel, per antenna, complex gain which flattens the bandpass and removes small phase slopes due to variation in cable delay. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.  These weights translate into a convolution in the delay space by a ringing harmonic point spread function.  We deconvolve using a one dimensional CLEAN algorithm   \citep{tay99} as described for the delay axis by \citet{par09,par12}.

@@

We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}. Boundaries of the {\it foreground wedge} are also shown. The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries. This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher. Emission in {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.

@@

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly. This is predominant at short antenna spacings. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions are valid despite imperfections in deconvolution.

>>>>

\subsubsection{Initial Data processing}\label{sec:data-analysis}
The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference, removing 3\% of the data and averaged in time  and frequency from the raw 1/2s,40kHz to 2s, 80khz. These data are then calibrated to a simulation of the sky containing 2420 point sources selected from the MWA Commissioning Survey (N.H Walker, in prep).  MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.        The calibration sources are selected to fall within the 5\% primary lobe of the tile beam. The calibration algorithm computes a per channel, per antenna, complex gain solutions averaged to two minute intervals.  The solutions are still fairly low signal to noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna to antenna variation. First we average the channel gains over all antenna to obtain a high signal to noise measurement of the bandpass.  After applying this single passband, we do a second round of calibration and fit a 2nd order amplitude and first order phase for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay. Finally we fit for an additional phase known to be caused by reflection in 150m cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.  In delay space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).  We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.   We limit the selection of peaks to modes inside the delay horizon corresponding to smooth spectrum sources on the sky --ie beneath the black line in Figure \ref{fig:fhd_data}.

@@

We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}. In this plot and others, we plot the amplitude of the delay spectrum sorted by baseline length. Note that no baselines have been averaged together in this plot.  The principle features are the wedge-shaped feature the boundary of which increases approximately linearly with light travel time of the baseline length.  Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different sources in the sky. It is this variation that we are investigating.

In these power spectrum units [Mpc$^{-3}$], the reionization power spectrum is expected to be spherically symmetric and decrease quickly with $|k|$.  The brightest signal will occur on the shortest baselines and smallest delays. Thus the region of interest is just above the black horizon line and below the harmonic due to the coarse channels.

 The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries. This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher. Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.

@@

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly. This is predominant at short antenna spacings where the number of delay modes within the horizon. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions are valid despite imperfections in deconvolution.
